---
title: "Portfolio Simulations"
author: "Bryan Bruno"
date: "09-28-2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Hedge funds can earn high proﬁts through the use of leverage, but leverage also creates high risk. The simulations in this section explore the eﬀects of leverage in a simpliﬁed setting.

Suppose a hedge fund owns $1,000,000 of stock and used $50,000 of its own capital and $950,000 in borrowed money for the purchase. Suppose that if the value of the stock falls below $950,000 at the end of any trading day, then the hedge fund will sell all the stock and repay the loan. This will wipe out its $50,000 investment. The hedge fund is said to be leveraged 20:1 since its position is 20 times the amount of its own capital invested. Suppose that the daily log returns on the stock have a mean of 0.05/year and a standard deviation of 0.23/year. These can be converted to rates per trading day by dividing by 253 and √253, respectively.

**What is the probability that the value of the stock will be below $950,000 at the close of at least one of the next 45 trading days?**


```{r problem 4}
niter = 1e5 # number of iterations
below = rep(0, niter) # set up storage

set.seed(2009)

for (i in 1:niter){
  r = rnorm(45, mean = 0.05/253, sd = 0.23/sqrt(253)) # generate random numbers
  logPrice = log(1e6) + cumsum(r)
  minlogP = min(logPrice) # minimum price over next 45 days
  below[i] = as.numeric(minlogP < log(950000))
  }
mean(below)
```

Suppose the hedge fund will sell the stock for a proﬁt of at least $100,000 if the value of the stock rises to at least $1,100,000 at the end of one of the ﬁrst 100 trading days, sell it for a loss if the value falls below $950,000 at the end of one of the ﬁrst 100 trading days, or sell after 100 trading days if the closing price has stayed between $950,000 and $1,100,000.

**What is the probability that the hedge fund will make a proﬁt of at least $100,000?**

**What is the probability the hedge fund will suﬀer a loss?**

**What is the expected proﬁt from this trading strategy?**

```{r problem 5, 6, 7}
# iterations
niter = 1e5 
sim_days = 100

# storage variables
below = rep(0, niter) 
above = rep(0, niter) 
middle = rep(0, niter)
profit = rep(0, niter)

# thresholds
prof_thresh = 1e6 + 1e5
log_prof_thresh = log(prof_thresh)
loss_thresh = 1e6 - 5e4
log_loss_thresh = log(loss_thresh)

set.seed(2009)

for (i in 1:niter){
  r = rnorm(sim_days, mean = 0.05/253, sd = 0.23/sqrt(253)) 
  logPrice = log(1e6) + cumsum(r)

  minlogP = min(logPrice)
  maxlogP = max(logPrice)
  
  p_thresh_hit = maxlogP >= log_prof_thresh
  l_thresh_hit = minlogP <= log_loss_thresh
  
  if(p_thresh_hit & !l_thresh_hit){
    above[i] = 1
    profit[i] = +1e5
  }

  if(!p_thresh_hit & l_thresh_hit){
    below[i] = 1
    profit[i] = -5e4
  }
    
  if(p_thresh_hit & l_thresh_hit){
    days_max = which(logPrice >= log_prof_thresh)[1]
    days_min = which(logPrice <= log_loss_thresh)[1]
    
    if(days_min < days_max){
      below[i] = 1
      profit[i] = -5e4
    }
    
    else{
      above[i] = 1
      profit[i] = +1e5
    }
  }

  if(!p_thresh_hit & !l_thresh_hit){
    middle[i] = 1
    profit[i] = exp(logPrice[sim_days]) - 1e6
  }
}
print("Solutions:")
print(round(mean(above)*100, 3))
print(round(mean(below)*100, 3))
print(round(mean(profit), 2))
```

**There is a 38.78% probability that the hedge fund will make a profit of at least $100,000.**

**There is a 58.84% probability that the hedge fund will suffer a loss.**

**The expected profit from this trading strategy is $9922.63.**
