---
title: "Stock Returns & Lognormal Distributions"
author: "Bryan Bruno"
date: "09-30-2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Part 1

Assume that a stock's log returns at any time scale have normal distribution. Suppose that its average annual log return is 100%, and its annual standard deviation ("volatility") of log returns is 200%. What are its average (mu) and standard deviation (sigma) of daily log returns, assuming a year has 250 trading days?

```{r part 1}
# trading days
t = 250

1.00/t
2.00/sqrt(t)
```
**The average (mu) of daily log returns is 0.004 for 250 trading days.**

**The standard deviation (sigma) of daily log returns is 0.1264911 for trading days.**

### Part 2

Simulate 250 instances of the daily log returns described in 1) with random seed set.seed(2015). Compute the net returns of these instances, and compute their average and standard deviation. Are the average (mu) and standard deviation (sigma) of net returns same as the average and standard deviation of log returns computed in 1)?

```{r part 2}
# seed set
set.seed(2015)

# trading days
t = 250

# simulation
sim = rnorm(t, 1.00/t, 2.00/sqrt(t))

# net return
net = exp(sim) - 1
  
m = mean(net)
s = sd(net)

m
s
```
**The average (mu) of this simulation's net returns is 0.006946427.**

**The standard deviation (sigma) of this simulation's net returns is 0.1210847.**

**While not exact, both the mu and sigma results are similar to Part 1 and within a reasonable range. Simulations will differ along with changing the random seed value. However, these results are relatively accurate given the number of simulation days.**

### Part 3.1

Compute the quantity m - (s^2)/2. How does this compare with the average log return mu computed in part 1)?  

```{r part 3.1}
mu = m - (s**2)/2
mu
```
**The quantity m - (s^2)/2 is -0.0003843233. At first glance, this may appear off from the average log return mu in Part 1; however, that is not actually the case. 250 is a very small sample size and the random seed also plays a factor as well. Given a large enough sample size, it will converge to normality.**

### Part 3.2

Their equality can be proven analytically through a mathematical theorem called Ito's Lemma that lies at the foundation of Black-Scholes options pricing formula. Their numerical equality is not that good here because Ito's Lemma assumes we can divide a period into infinitesimally small sub-periods. So divide a year into 25,000 sub-periods (think of these sub-periods roughly as minutes) instead, and compare the new mu (average log return per minute) with the new m - (s^2)/2 (m is now the average net return per minute). Also compare the new sigma (standard deviation of log returns per minute) with the new s (standard deviation of net return per minute).

```{r part 3.2}
# seed set
set.seed(2015)

# trading seconds
t_s = 25000
mu_s = 1.00/t_s
sig_s = 2.00/sqrt(t_s)

# simulation
sim_s = rnorm((t * 6.5 * 60), mu_s, sig_s)

# net return
net_s = exp(sim_s) - 1

m_s = mean(net_s)
s_s = sd(net_s) # (4) new sd of net return/min
s_s2 = sd(sim_s) # (3) new sd of log return/min
mu_s2 = format(mean(sim_s), scientific = FALSE) # (1) new ave log return/min
mu_2 = format(m_s - (s_s**2)/2, scientific = FALSE) # (2) new m - (s^2)/2

mu_s2
mu_2
s_s2
s_s
```
**The new mu (average log return per minute) of this simulation is 0.00009285604.**

**The new m - (s^2)/2 (average net returns per minute) of this simulation is 0.00009282939.**

**The new sigma (sd of log returns per minute) of this simulation is 0.01267206.**

**The new s (sd of net returns per minute) of this simulation is 0.01267554.**

### Part 4

If we assume that the stock's initial price is $1, what is the expected value of its log price log(P(t)) after t minutes expressed in terms of mu? And what is the expected value of its price P(t) expressed in terms of mu and sigma? Finally, express these expected values in terms of mu and sigma instead.


**To first find the log price log(P(t)) after t minutes, we know the non-log price can be expressed by t * (m - (s^2)/2). With the following key concept in mind, E[exp(x)] != exp(E[x]), we need to derive mu and sigma correctly.**

**If Z is distributed N(mu, sigma^2), then Y = exp(Z) is said to have a Lognormal(mu, sigma^2) distribution.**

**Thus, E[exp(x)] = exp((mu + sigma^2)/2).**

**Further, mu = t * (m - (s^2)/2) and sigma^2 = t * s^2.**