---
title: "Fitting Bivariate T-Distributions"
author: "Bryan Bruno"
date: "10-8-2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Setup Code

```{r code}
library(mnormt)

data(CRSPday, package = "Ecdat")

Y = CRSPday[ , c(5, 7)]

loglik = function(par){
  mu = par[1:2]
  A = matrix(c(par[3], par[4], 0, par[5]), nrow = 2, byrow = T)
  scale_matrix = t(A) %*% A
  df = par[6]
  -sum(log(dmt(Y, mean = mu, S = scale_matrix, df = df)))
}

A = chol(cov(Y))
start = as.vector(c(apply(Y, 2, mean), A[1, 1], A[1, 2], A[2, 2], 4))
fit_mvt = optim(start, loglik, method = "L-BFGS-B",
                lower = c(-0.02, -0.02, -0.1, -0.1, -0.1, 2),
                upper = c(0.02, 0.02, 0.1, 0.1, 0.1, 15), hessian = T)
```

### Part A
What does the code A = chol(cov(Y)) do?

```{r code_look}
head(Y)
print(cov(Y))
print(A)
```
**To put this equation into plain English, A is equal to the Cholesky factorization of the covariance matrix of Y. In this case, Y contains the stock returns of both IBM and CRSP. In the code above, I have outlined the head of Y, which displays the first six returns of IBM and CRSP. Next, I show the covariance 2 by 2 matrix of Y. Finally, I output the resulting Cholesky factorization of the matrix, A.**

### Part B
Find the MLE of θ.

```{r part_b}
fit_mvt$par
```

### Part C
Find the Fisher information matrix for θ.

```{r part_c}
fit_mvt$hessian
fisherinv = solve(fit_mvt$hessian, diag(length(fit_mvt$par)))
fisherinv
```

### Part D
Find the standard errors of the components of the MLE of θ using the Fisher information matrix.

```{r part_d}
sqrt(diag(fisherinv))
```

### Part E
Find the MLE of the covariance matrix of the returns.

```{r part_e}
parfit = fit_mvt$par
R = matrix(c(parfit[3], parfit[4], 0, parfit[5]), nrow = 2, byrow = T)
covmat = t(R) %*% R
covmat
```

### Part F
Find the MLE of ρ, the correlation between the two returns (Y1 and Y2).

```{r part_f}
YD = diag(c(sqrt(covmat[1, 1]), sqrt(covmat[2, 2])))
invYD = solve(YD, diag(2))
cormat = invYD %*% (covmat %*% invYD)
cormat
```